Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2021-10-26 analysis/paper_draft2.Rmd
revised text of Discussion

Statistical methods for _Batson_ and similar changes are intended to measure how much prospective jurors' race, gender, or other cognizable-class membership matters to how likely an attorney will use peremptory strikes against them. The Bayesian approach here accounts for the extent of the available historical strike data when estimating the posterior for the bias parameter. By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data.

And as our prototype app shows, if the historical strike data loaded into the app in advance, attorneys and others can the approach here in real time during jury selection. Despite _Batson_ and similar law, racial bias persists in the use of peremptory challenges [@Diamond2018], perhaps in part because attorneys fear that _raising_ a _Batson_ challenge raises with it the embarrassment and discomfort associated with accusing another attorney of being racist [@OBrien2019, 29; @Offit2021]. It is worth studying whether an app like the prototype presented here might reduce that and other barriers to bringing _Batson_ challenges.

@@

Finally, for the approach here, like any bias-detection method relying on historical strike data, the ideal data is, for each attorney, a complete history of who they struck in all their past cases. The problem: Jury-selection records in past cases are hard to get because of court practices and other barriers [@Grosso2017; @Wright2018]. For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender). This forces researchers to turn to other data sources (e.g., voter registration rolls), if any, to find that information. And judges and court staff vary in how often they fully complete or maintain the information they are required to record. Incomplete historical strike data resents an issue for inferring bias if such data is incomplete or missing not at random. These data challenges, however, might decrease over time as judges, attorneys, and jury clerks realize that more historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.

>>>>

Statistical methods for _Batson_ and similar changes aim to help measure how much prospective jurors' race, gender, or other cognizable-class membership matters to how likely an attorney will use peremptory strikes against them. The Bayesian approach here accounts for the extent of the available historical strike data when estimating the posterior for the bias parameter. By incorporating the power prior, one can easily assess how sensitive that posterior estimate is to the available historical strike data. And with a tool like our prototype app, attorneys can use historical strike data to detect bias in real time during jury selection, thus making it easier to raise credible _Batson_ challenges.

@@

Finally, for the approach here, incomplete historical strike data presents an issue for inferring bias if such data is missing not at random. In practice, jury-selection records in past cases are hard to get because of court practices and other barriers [@Grosso2017; @Wright2018]. For example, some jurisdictions do not require recording relevant or detailed enough information about prospective jurors who were or could have been struck (e.g., their race or gender). This forces researchers to turn to other data sources, if any, to find that information. And judges and court staff vary in how often they fully complete or maintain the information they are required to record. We hope these data challenges decrease over time as judges, attorneys, and jury clerks realize that more historical strike data makes it easier, with the approach described here, to credibly infer or disprove bias under _Batson_ and similar law.
